A Neural Network Alternative to Convolutive Audio Models for Source Separation

### 2017-06-13 paper/paper.tex
I have made few more changes on the paper.

\section{Experiments}
\label{sec:experiments}
We now describe the experimental setup used to evaluate our auto-encoder based convolutive audio models. We construct a set of training and test examples using the TIMIT corpus~\cite{timit} for the evaluation. To form the examples, we randomly select a pair of male-female speakers from the TIMIT corpus. Of the $10$ utterances available for each speaker, one utterance is randomly selected for each speaker. These two selected utterances are mixed at $0~dB$ to generate the testing mixture. The remaining $9$ utterances are used as training data to construct models for the sources. In other words, these examples are used to train the respective (convolutional/ feed-forward) auto-encoders. For the evaluation, we generate $20$ such mixtures and compare the models for different parameter configurations. As a pre-processing step, we apply a $1024$ point short-time Fourier transform representation with a hop of $25\%$. The magnitude spectrogram is then given as an input to the network. \\

The neural networks are initialized using the Xavier initialization scheme~\cite{glorot2010understanding}. The networks are trained by applying a batch gradient descent training procedure and the parameters updated using the RMSProp algorithm~\cite{tieleman2012rmsprop}, with a learning rate and momentum of $0.001$ and $0.7$ respectively. \\

The CNN filters are selected to be $512$-point tall and $8$-point wide. Thus, the convolutions are performed only along the time axis. The number of CNN filters also decides the number of components in the decomposition. We evaluate these models over a varying number of CNN filters ranging from $10$ to $100$ uniformly in steps of $10$. We compare the separation performance in terms of median BSS\_eval metrics~\cite{fevotte2005bss_eval} viz., signal-to-distortion (SDR), signal-to-interference (SIR) and signal-to-artifact ratio (SAR) parameters. \\

\subsection{Results and Discussion}
\label{subsec:results}
Figure~\ref{fig:cnnseparation_performance} gives the separation performance for the CNN-CNN models (green) and RNN-CNN models (red) for varying values of number of filters~$K$. We evaluate the proposed CAE models by comparing the separation performance to their equivalent feed-forward~(FF) counterparts (blue) proposed in~\cite{smaragdis2017aneural}. In order to maintain a uniform experimental setup, we apply the separation scheme described in~\ref{sec:ss} to all the models. We plot the results in terms of a violin plot that also indicates the median value (dotted line in the centre) and the corresponding inter-quartile range (outer dotted lines) for each $K$. \\

@@

\section{Conclusion}
\label{sec:conclusion}
In this paper, we developed and investigated the use of a convolutional auto-encoder as an alternative to learn convolutive basis decompositions of speech and audio signals. The ability of the networks to include temporal dependencies allow the auto-encoders to learn cross-frame structures in the input spectrogram. We demonstrated that this results in a significant improvement in separation performance as compared to feed-forward auto-encoder models. This approach also allows for several extensions and generalizations to convolutive audio models, that can be easily implemented using the available diversity of neural networks. One such extension considered in this paper is the use of auto-encoders formed by a cascade of recurrent and convolutive layers. Although these models are not as potent as CAE models for separation of speech mixtures, these extensions could lead to interesting non-uniform auto-encoder architectures for several applications. \\

>>>>

\section{Experiments}
\label{sec:experiments}
We now describe the experimental setup used to evaluate our auto-encoder based convolutive audio models. We construct a set of training and test examples using the TIMIT corpus~\cite{timit} for the evaluation. To form the examples, we randomly select a pair of male-female speakers from the TIMIT corpus. Of the $10$ utterances available for each speaker, one utterance is randomly selected for each speaker. These two selected utterances are mixed at $0~dB$ to generate the testing mixture. The remaining $9$ utterances are used as training data to construct models for the sources. In other words, these examples are used to train the respective (convolutional/ feed-forward) auto-encoders. For the evaluation, we generate $20$ such mixtures and compare the models for different parameter configurations. As a pre-processing step, we apply a $1024$ point short-time Fourier transform representation with a hop size of $25\%$. The magnitude spectrogram is then given as an input to the network. \\

The networks are trained by applying a batch gradient descent training procedure and the parameters updated using the RMSProp algorithm~\cite{tieleman2012rmsprop}, with a learning rate and momentum of $0.001$ and $0.7$ respectively. The neural networks are initialized using the Xavier initialization scheme~\cite{glorot2010understanding}.\\

The CNN filters are selected to be $512$-point tall and $8$-point wide. The number of CNN filters also decides the number of components in the decomposition. We evaluate these models over a varying number of CNN filters ranging from $10$ to $100$ uniformly in steps of $10$. We compare the separation performance in terms of median BSS\_eval metrics~\cite{fevotte2005bss_eval} viz., signal-to-distortion (SDR), signal-to-interference (SIR) and signal-to-artifact ratio (SAR) parameters. \\

\subsection{Results and Discussion}
\label{subsec:results}
Figure~\ref{fig:cnnseparation_performance} gives the separation performance for the CNN-CNN models (green) and RNN-CNN models (red) for varying values of number of filters~$K$. We evaluate the proposed CAE models by comparing the separation performance to their equivalent feed-forward~(FF) counterparts (blue) proposed in~\cite{smaragdis2017aneural}. In order to maintain a uniform experimental setup, we apply the separation scheme described in~\ref{sec:ss} to all the models. We plot the results in terms of a violin plot that also indicates the median value (dotted line in the center) and the corresponding inter-quartile range (outer dotted lines) for each $K$. \\

@@

\section{Conclusions}
\label{sec:conclusion}
In this paper, we developed and investigated the use of a convolutional auto-encoder as an alternative to learn convolutive basis decompositions of speech and audio signals. The ability of the networks to include temporal dependencies allow the auto-encoders to learn cross-frame structures in the input spectrogram. We demonstrated that this results in a significant improvement in separation performance as compared to feed-forward auto-encoder models. This approach also allows for several extensions and generalizations to convolutive audio models, that can be easily implemented using the available diversity of neural networks. One such extension considered in this paper is the use of auto-encoders formed by a cascade of recurrent and convolutive layers. Although this extension have not outperformed the convolutive architecture, we have shown that this architecture outperforms the standard feedforward network. \\
